DRAFT: Harmonized and trait aggregation paper

### 2020-10-16 Paper/Draft_harmonization_paper.tex
16.10: Paper Update, first implm. of Co-Authors

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{hyperref}
\usepackage{float}
\usepackage[table,xcdraw]{xcolor}
\usepackage{color, colortbl}
\usepackage{longtable}
%\usepackage[sort&compress,square,comma,authoryear]{natbib}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{tikz}
% \graphicspath{{/home/kunz/Dokumente/Projects/Trait_DB/Invertebrate_traits/Paper/Figures/}}
\graphicspath{{Figures/}}
\usepackage{rotating}
\usepackage{geometry}
\usepackage{array}
\usepackage{lscape}
\usepackage{longtable}
\usepackage{hhline}
\usepackage[
backend = biber,
style = ieee,
citestyle = numeric-comp,
maxbibnames=99,
sorting = none % sort by name year title
]{biblatex}
\addbibresource{Ref_invertebrate_DB.bib}

@@

Explaining and predicting how communities are shaped by environmental factors is one of the main goals of ecology. Species traits, defined as measurable properties of an organism \cite{mcgill_rebuilding_2006}, might be beneficial in achieving this goal \cite{heino_jani_macroecological_2013}. %?just traits 
Traits evolve through adaptations (e.g., physiological, behavioural, etc.) of organisms to their environment and indicate direct or indirect linkages between the biological response of an organism to its environment \cite{southwood_habitat_1977, verberk_delivering_2013}.
% TODO: PUP: Not only „of an organism“, but also „of a population“ (at least for some of the traits related to reproduction; e.g. difficult to describe „voltinism“ as the organism scale!).
Besides providing a mechanistic explanation of species-environment relationships, trait-based approaches might be suitable for large scale analysis because the variability in trait responses is lower than in taxonomic responses \cite{bonada_taxonomic_2007, baird_toward_2011}. Traits of freshwater invertebrate individuals are difficult to determine because - unlike plants - their traits often cannot be measured directly.
For example, to gain knowledge on feeding habits requires evaluating mouthpart morphology, consumed food, and the organisms function within its community \cite{moog_comprehensive_nodate}. 
% PUP says: But the trait-based approach can benefit from two centuries of detailed, published observations on biological traits and ecological preferences made by zoologists/ecologists on a large range of species in various environmental conditions.
Nevertheless, invertebrate traits have been increasingly used in freshwater ecology, e.g. by relating macroinvertebrate trait composition to environmental factors or as trait metrics in biomonitoring \cite{poff_developing_2010, szocs_effects_2014, bhowmik_large_2015, menezes_beyond_2010}.

In the last decades, freshwater ecologists developed comprehensive invertebrate trait databases for various biogeographic regions \cite{usseglio-polatera_biomonitoring_2000, schmidt-kloiber_www.freshwaterecology.info_2015, vieira_database_nodate, Philips_and_Smith_NZ_DB_2018, kefford_integrated_2020, tomanova_trophic_2006}. The availability of invertebrate trait data from different biogeographic regions enables comparisons of trait variation and their relation to environmental factors across large scales. However, such analyses have been carried out mostly within biogeographic regions, using information from one or two trait databases. For example, Bonada et al. \cite{bonada_taxonomic_2007} compared trait composition for Mediterranean and temperate regions in Europe using traits from Usseglio-Polatera et al. \cite{usseglio-polatera_biomonitoring_2000}, Poff et al. \cite{poff_developing_2010} characterized trait composition across sites in the Western US using traits from Poff et al. \cite{poff_functional_2006}, and Botwe et al. \cite{botwe_effects_2018} investigated the effect of salinity on invertebrate traits in different sites in South Australia using trait data from Poff et al. \cite{poff_functional_2006} and Schäfer et al. \cite{schafer_trait_2011}. 
% ? Include Saito for South America 
Analyses of invertebrate traits that synthesize information on invertebrate grouping features from more than two different biogeographic regions are rare. A grouping feature is defined by Schmera et al. as a general property (e.g. feeding mode) that comprises a "group of related traits (e.g., predator, shredder, etc.) that vary among species or among individuals within a species" \cite{schmera_proposed_2015}. To our knowledge, only Brown et al. \cite{brown_functional_2018} harmonized grouping features from more than two geographically distant invertebrate trait databases in a study on the influence of decreasing glacier cover on functional diversity and community assembly of invertebrates. 
% TODO: ?Paper trait convergence Doledec

We suspect that the heterogeneity of information in freshwater invertebrate trait databases is likely a major reason for the lack of studies across biogeographic regions. To harmonize grouping features from different regions, first commonly accepted and unambiguous trait definitions are required \cite{schneider_towards_2019}. In the best case, grouping features would be classified into the same traits across databases or they could easily be harmonized using standardized terminology. However, a lack of standardized terminology of trait definitions and poor metadata quality in many trait databases is a common issue throughout the field of trait-based ecology \cite{baird_toward_2011, schneider_towards_2019}. Secondly, consistent coding of traits facilitates the compatibility of trait data from different databases. Traits can be binary (i.e. trait exists or not), continuous, or fuzzy coded variables. Fuzzy codes represent the affinity of an organism to express a certain trait. They are used to account for plasticity in traits and are usually converted to percentages. Continuous values are typically used for grouping features that can be measured, like body size. However, invertebrate trait databases are heterogeneous with regard to the coding they use for their traits \cite{culp_incorporating_2011} (Table \ref{tab:trait_databases_coding_differentiation}). Brown et al. \cite{brown_functional_2018} harmonized grouping features based on trait databases from Europe, North America, and New Zealand because in these trait databases identical grouping features are classified differently into traits. As the traits from North America were coded binary in contrast to the fuzzy coded traits from Europe and New Zealand, the authors consulted experts to assign fuzzy coded traits to North American taxa or inferred them from the European trait database. Thus, it becomes apparent that using invertebrate trait data from several regions requires extensive data processing prior to the actual data analysis. A centralized database with standardized and unambiguous traits and a consistent coding of traits would minimize data processing effort.

Differing taxonomic resolutions between the observed taxa in a study and the used trait database is another challenge when working with trait data. When observations are on a lower taxonomic level than data available in the trait databases (e.g. observations on species-level, trait data on genus-level) trait data of the higher taxonomic level are often assigned, e.g. \cite{szocs_effects_2014, vos_taxonomic_2017}. Conversely, if trait information is only available on lower taxonomic levels than the observed taxa, traits are aggregated to a higher taxonomic level, e.g. \cite{poff_functional_2006, szocs_effects_2014, piliere_a._f._h._importance_2016, aspin_extreme_2019}. Thereby, trait aggregation is often done using the mean \cite{magliozzi_functional_2019}, median \cite{szocs_effects_2014} or the mode \cite{piliere_a._f._h._importance_2016}. Up to now, studies on how and to which extent different trait aggregation methods influence trait-based analysis are missing. Knowing this can aid the choice of aggregation method for future studies.

We analyzed the influence of grouping feature harmonization and trait aggregation on trait-environment relationships. 
Therefore, we (1) harmonized four invertebrate grouping feature datasets for seven grouping features based on information from trait databases of the regions Europe, North America, New Zealand, and Australia. Furthermore, we (2) compare trait affinities obtained through different trait aggregation methods to trait affinities assigned at family-level by experts. We (3) re-analyze data on the effect of anthropogenic salinization on biological traits by Szöcs et al. \cite{szocs_effects_2014} using harmonized grouping features and aggregated traits. By comparison with the original analysis, we investigate how harmonizing and aggregating trait data can alter the outcome of trait-environment relationships. Finally, we (4) present an overview of discrepancies in trait definition between the used invertebrate trait datasets and discuss challenges of trait data synthesis.

@@

We extracted information from 6 trait databases from Europe, North America, Australia, and New Zealand and harmonized 7 grouping features. Trait information for Europe was obtained from the Freshwaterecology database \cite{schmidt-kloiber_www.freshwaterecology.info_2015} and complemented by Tachet \cite{usseglio-polatera_biomonitoring_2000} to fill missing information (e.g. for size).
% TODO: mention that tachet and freshwaterecology are not seperate (but we did it separately)
Trait information for North America was obtained from Twardochleb et al. \cite{twardochleb_trait_data_2020} and complemented by Vieira et al. \cite{vieira_database_nodate}. Data on body form for European and North American taxa was provided based on expert knowledge \cite{polatera_personal_information_2020}. For Australia and New Zealand, we used trait databases from Kefford et al. \cite{kefford_integrated_2020} and Philips and Smith respectively \cite{Philips_and_Smith_NZ_DB_2018}.

We selected traits of seven grouping features that were available in all databases, are commonly used, and describe different parts of the biology of an organism: life history (Voltinism), morphology (Respiration, Body form, Size), ecology (Locomotion, Feeding mode) and reproduction (Oviposition). We omitted ecological traits that describe habitat preferences (e.g. temperature preference) because these traits are missing in the New Zealand trait database. The grouping features were differently classified across the databases, we therefore harmonized them into 26 traits (Table \ref{tab:traits_harmonization}). Harmonization was undertaken by amalgamating similar traits into one trait (e.g. crawlers and sprawlers into crawlers). Thereby, the highest trait affinity among the amalgamated traits was taken.

We used fuzzy coded traits for establishing our harmonized datasets unless data quality prohibited and then we used binary traits, i.e. categorical and continuous traits were converted into binary traits. Implicitly, we assumed for binary traits that a value of 1 corresponds to the highest affinity for a particular trait and 0 to no affinity for a particular trait. Fuzzy codes are reported with different ranges in the trait databases (e.g. freshwaterecology 0 to 10, Tachet 0 to 3 or 0 to 5). We standardized them to a range between 0 and 1 and converted trait affinities to percentages. Thus, fuzzy coded and binary traits were in the same range.

Prior harmonization we amalgamated duplicate taxa on species, genus or family-level if present. We omitted taxa with a lower taxonomic resolution than family-level.

\begin{table}[H]
\centering
\caption{Traits of the harmonized grouping features. The last column indicates traits that were amalgamated for harmonization (no amalgamation needed if empty).}
\label{tab:traits_harmonization}
\begin{tabular}{lll}
\toprule[.1em]
Grouping feature & Trait & Amalgamated traits\\
\toprule[.1em]
Voltinism    & \begin{tabular}[c]{@{}l@{}}Semivoltine\\ Univoltine\\ Bi/multivoltine\end{tabular}                                & \begin{tabular}[c]{@{}l@{}}\textless 1 generation per year\\ 1 generation per year\\ \textgreater 1 generation per year\end{tabular}                                                                                                                                            \\
\midrule
Body Form    & \begin{tabular}[c]{@{}l@{}}Cylindrical \\ Flattenend\\ Spherical\\ Streamlined\end{tabular}                       & \begin{tabular}[c]{@{}l@{}}Cylindrical, round (humped)\\ Flattend, dorsoventrally flattened\\ Spherical, tubular\\ Streamlined, fusiform\end{tabular}                                                                                                                                                                          \\
\midrule
Size         & \begin{tabular}[c]{@{}l@{}}Small \\ Medium \\ Large\end{tabular}                                                  & \begin{tabular}[c]{@{}l@{}}\textless 9 mm, \textless 10 mm\\ 9 - 16 mm, 10 - 20 mm\\ \textgreater 16 mm, \textgreater 20 mm\end{tabular}                                                                                                                \\
\midrule
Respiration  & \begin{tabular}[c]{@{}l@{}}Gills\\ Plastron/Spiracle\\ \\ \\ Tegument\end{tabular}                             & \begin{tabular}[c]{@{}l@{}}Temporary air store, tracheal gills, gills\\ Spiracular gills, atmospheric breathers, \\ plant breathers, functional spiracles, \\ air (plants), aerial, plastron/spiracle\\ Cutaneous, tegument \end{tabular}                                                                         \\
\midrule
Locomotion   & \begin{tabular}[c]{@{}l@{}}Burrower\\ Crawler\\ Sessil\\ Swimmer \end{tabular}                                     & \begin{tabular}[c]{@{}l@{}}Interstitial, boring, burrowing\\ Sprawler, walking, climber, clinger, crawler\\ Attached, sessil\\ Skating, diving, planctonic, swimming\end{tabular}                                                                                                                 \\
\midrule
Feeding mode & \begin{tabular}[c]{@{}l@{}}Filterer\\ \\ Gatherer\\ \\ Herbivore\\ \\ Parasite\\ Predator\\ Shredder \\ \\ \end{tabular} & \begin{tabular}[c]{@{}l@{}}Active/passive filterer, absorber, \\ filter-feeder, collector-filterer, filterer\\ Deposit-feeder, collector-gatherer, \\ detrivore, gatherer\\ Grazer, scraper, piercer herbivore, \\ herbivore, algal piercer, piercer (plants)\textsuperscript{\textit{a}}\\ \\ Piercer (animals)\textsuperscript{\textit{a}}, predator \\ Miner, xylophagus, shredder, \\ shredder detrivore\end{tabular} \\
\hline
Oviposition  & \begin{tabular}[c]{@{}l@{}}Aquatic eggs\\ \\ Ovoviviparity\\ Terrestrial eggs\end{tabular}                            & \begin{tabular}[c]{@{}l@{}}Eggs attached to substrate/plants/stones,\\ free/fixed eggs/clutches\\ \\ Terrestrial clutches, terrestrial \end{tabular}     \\
\bottomrule[.1em]
\end{tabular}
\end{table}
\begin{minipage}{\linewidth}\small
    \textit{a} The trait piercer was defined in the Tachet database for piercing plants and animals, in contrast to the other databases \cite{usseglio-polatera_biomonitoring_2000}. Taxa exhibiting this trait have been assigned to predators or herbivores based on expert knowledge \cite{polatera_personal_information_piercer_2020}.
\end{minipage}

@@

Traits of the harmonized grouping feature datasets were aggregated to family-level using three approaches. I) direct aggregation of taxa to family-level giving equal weight to every species using the mean or median, denoted \textit{direct\_agg\textsubscript{mean}} and \textit{direct\_agg\textsubscript{median}}, respectively. II) stepwise aggregation, i.e. first to the genus-level and subsequently to the family-level using the mean or median. This approach gives equal weights to each genus. Hereafter, we denote this aggregation type as \textit{stepwise\_agg\textsubscript{mean}} or \textit{stepwise\_agg\textsubscript{median}}, respectively. III) aggregation using a weighted mean approach, denoted as \textit{weighted\_agg}. This method weights the genera according to the number of their species in the databases.

\begin{figure}
  \centering
  \subfile{Flowchart/Flowchart_methods.tex}
  \caption{Data processing steps of the selected traits. Intermediate (gray) and main (orange) steps of data preparation are depicted. The dashed bottom box illustrates the different trait aggregation methods using a small made-up example (data in the upper left corner). Displayed are the aggregation methods (blue) and intermediate steps of the aggregation methods (purple). For the stepwise and weighted aggregation only results for one trait are displayed.}
\end{figure}

>>>>

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{hyperref}
\usepackage{float}
\usepackage[table,xcdraw]{xcolor}
\usepackage{color, colortbl}
\usepackage{longtable}
%\usepackage[sort&compress,square,comma,authoryear]{natbib}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{tikz}
% \graphicspath{{/home/kunz/Dokumente/Projects/Trait_DB/Invertebrate_traits/Paper/Figures/}}
\graphicspath{{Figures/}}
\usepackage{rotating}
\usepackage{geometry}
\usepackage{array}
\usepackage{lscape}
\usepackage{longtable}
\usepackage{hhline}
\usepackage{lmodern}
\usepackage[
backend = biber,
style = ieee,
citestyle = numeric-comp,
maxbibnames=99,
sorting = none % sort by name year title
]{biblatex}
\addbibresource{Ref_invertebrate_DB.bib}

@@

Explaining and predicting how communities are shaped by environmental factors is one of the main goals of ecology. Species traits, defined as measurable properties of an organism \cite{mcgill_rebuilding_2006}, might be beneficial in achieving this goal \cite{heino_jani_macroecological_2013}. %?just traits 
Traits evolve through adaptations (e.g., physiological, behavioural, etc.) of organisms to their environment and indicate direct or indirect linkages between the biological response of an organism to its environment \cite{southwood_habitat_1977, verberk_delivering_2013}.
%%%%%%%%%%%%%%%%%%%%%%%%%%
% TODO: PUP: Not only „of an organism“, but also „of a population“ (at least for some of the traits related to reproduction; e.g. difficult to describe „voltinism“ as the organism scale!).
%%%%%%%%%%%%%%%%%%%%%%%%%%
Besides providing a mechanistic explanation of species-environment relationships, trait-based approaches might be suitable for large scale analysis because the variability in trait responses is lower than in taxonomic responses \cite{bonada_taxonomic_2007, baird_toward_2011}. Traits of freshwater invertebrate individuals are difficult to determine because - unlike plants - their traits often cannot be measured directly.
For example, to gain knowledge on feeding habits requires evaluating mouthpart morphology, consumed food, and the organisms function within its community \cite{moog_comprehensive_nodate}. 
%%%%%%%%%%%%%%%%%%%%%%%%%%
% PUP says: But the trait-based approach can benefit from two centuries of detailed, published observations on biological traits and ecological preferences made by zoologists/ecologists on a large range of species in various environmental conditions.
%%%%%%%%%%%%%%%%%%%%%%%%%%
Nevertheless, invertebrate traits have been increasingly used in freshwater ecology, e.g. by relating macroinvertebrate trait composition to environmental factors or as trait metrics in biomonitoring \cite{poff_developing_2010, szocs_effects_2014, bhowmik_large_2015, menezes_beyond_2010}.

In the last decades, freshwater ecologists developed comprehensive invertebrate trait databases for various continents \cite{usseglio-polatera_biomonitoring_2000, schmidt-kloiber_www.freshwaterecology.info_2015, vieira_database_nodate, Philips_and_Smith_NZ_DB_2018, kefford_integrated_2020, tomanova_trophic_2006}. The availability of invertebrate trait data from different continents enables comparisons of trait variation and their relation to environmental factors across large scales. However, such analyses have been carried out mostly within continents, using information from one or two trait databases. For example, Bonada et al. \cite{bonada_taxonomic_2007} compared trait composition for Mediterranean and temperate regions in Europe using traits from the Tachet database \cite{usseglio-polatera_biomonitoring_2000}, Poff et al. \cite{poff_developing_2010} characterized trait composition across sites in the Western US using traits from Poff et al. \cite{poff_functional_2006}, and Botwe et al. \cite{botwe_effects_2018} investigated the effect of salinity on invertebrate traits in different sites in South Australia using trait data from Poff et al. \cite{poff_functional_2006} and Schäfer et al. \cite{schafer_trait_2011}. Analyses of invertebrate traits that synthesize information on invertebrate grouping features from more than two different continents are rare. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% biogeographic regions substituted with continents
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% PUP & ASK: So „grouping features“ are „traits“ (in most of the published databases) described by „traits“ (i.e. several „related categories“). 
% I agree that these definitions have to be made more clear, i.e. that you are using the term „grouping feature“ for what most published articles call “traits” and that what others call “categories” (within the traits) are your “traits”. Maybe add a “definition” of terms at the beginning ...also your use of “affinity” needs to be included there…
% Put into definitons of terms?
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In this study, we follow the terminology proposed by Schmera et al., where a grouping feature is defined as a general property (e.g. feeding mode) that comprises a "group of related traits (e.g., predator, shredder, etc.) that vary among species or among individuals within a species" \cite{schmera_proposed_2015}. Thus, we use the term grouping feature in place of which many studies use the term "trait", and the term trait instead of "trait state", "modality" or "trait category". 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% TODO: Description of affinity
% A trait can described using a binary coding (exists or not), or using affinity scores that indicate the which extent an organism expresses a certain trait.  
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
To our knowledge, only Brown et al. \cite{brown_functional_2018} harmonized grouping features from more than two geographically distant invertebrate trait databases in a study on the influence of decreasing glacier cover on functional diversity and community assembly of invertebrates. 
% TODO: ?Paper trait convergence Doledec

We suspect that the heterogeneity of information in freshwater invertebrate trait databases, besides the diversity of taxa across regions, is likely a major reason for the lack of studies across continents. To harmonize grouping features from different regions, first commonly accepted and unambiguous trait definitions are required \cite{schneider_towards_2019}. In the best case, grouping features would be classified into the same traits across databases or they could easily be harmonized using standardized terminology. However, a lack of standardized terminology of trait definitions and poor metadata quality in many trait databases are common issues throughout the field of trait-based ecology \cite{baird_toward_2011, schneider_towards_2019}. Secondly, consistent coding of traits facilitates the compatibility of trait data from different databases. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% BK: I am not sure that the it is binary which is the defining issue, rather it is that there is no uncertainty. E.g. respiration type as more than two opitions but not undertainty into which a species fall into. 
% PUP: OK in your definition of „grouping features“ and „traits“, it is probably more the „trait description“ than the „trait“ (itself) which can be „binary“.
% ASK to Voltinism: I think this not an appropriate example for multiple categories except you include different regions for which different categories can apply; i.e. multivoltine in the temperate region and univoltine in the boreal region for the same species
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Alternative from Ben:
Traits can be described in a binary fashion with no uncertainty how the trait is expressed in any particularly organism (e.g. adult terrestrial stage, presence of gills), continuous (e.g. tolerance of pollution, body size) or using multiple descriptions (e.g. locomotion) with some uncertainty as to which trait the organism should be coded to. One approach for dealing with the uncertainty in the latter is the use of fuzzy coded variables. Fuzzy codes are used to account for plasticity in traits, variability in traits within taxonomic groups above species, and incomplete knowledge and are usually converted to proportions.

However, invertebrate trait databases are heterogeneous regarding the coding they use for their traits \cite{culp_incorporating_2011} 
%(Table \ref{tab:trait_databases_coding_differentiation}).
% TODO: Don't use table 5 and maybe find another reference?
Brown et al. \cite{brown_functional_2018} harmonized grouping features based on trait databases from Europe, North America, and New Zealand because in these trait databases identical grouping features are classified differently into traits. As the traits from North America were coded binary in contrast to the traits from Europe and New Zealand which have been established using affinity scores, the authors consulted experts to assign fuzzy coded traits to North American taxa or inferred them from the European trait database. Thus, it becomes apparent that using invertebrate trait data from several regions requires extensive data processing prior to the actual data analysis. A centralized database with standardized and unambiguous traits and a consistent coding of traits would minimize data processing effort.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% PUP: Taking into account your definition of traits (= „trait categories“ or „trait modalities“ in Tachet et al., 2010) and grouping features (= „traits“ in Tachet et al., 2010), the fuzzy coding approach is not applied to the „trait“ scale, but to the „grouping feature“ scale. At the „trait“ scale, „only“ an affinity score is allocated to a given taxon. The binary/quantitative character of this score is probably not the only important criterion to consider. One of the major characteristics of the fuzzy coding approach is, at the „grouping feature“ scale, to allow several „traits“ to be simultaneously present (possibly with different affinity scores) in a given „grouping feature“ for a given taxon (i.e. a non-disjunctive approach).
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ASK: this is something that should not appear in the introduction, at least not in such detail…
% ASK: I would suggest not to use “higher/lower” as this is used in different ways in different publications (low can mean low (i.e. imprecise) identification level, but it also can mean low in terms of systematic hierarchy, which then would mean more precise id level…); I would suggest to use terms like precise/imprecise, accurate or the like…
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Differing taxonomic resolutions (e.g. species, genus or family) between the observed taxa in a study and the used trait database is another challenge when working with trait data. When observations are on a more precise taxonomic level than data available in the trait databases (e.g. observations on species-level, trait data on genus-level) trait information of the less precise taxonomic level is often assigned (e.g. \cite{szocs_effects_2014, vos_taxonomic_2017}). Conversely, if trait information is only available on more precise taxonomic levels than the observed taxa, traits are aggregated to a less precise taxonomic level, e.g. \cite{poff_functional_2006, szocs_effects_2014, piliere_a._f._h._importance_2016, aspin_extreme_2019}. Thereby, trait aggregation is often done using the mean \cite{magliozzi_functional_2019},
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% PUP: Sometimes weighted by the rarity/commonality of individual taxa (e.g. genera) at large spatial scale (e.g. ecoregions), within the higher level taxonomic unit (e.g. family), because the probability of the commonest taxa to be present in a given sample is higher (at large spatial scale).
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
median \cite{szocs_effects_2014} or the mode \cite{piliere_a._f._h._importance_2016}. Up to now, studies on how and to which extent different trait aggregation methods influence trait-based analysis are missing. Knowing this can aid the choice of aggregation method for future studies.

We aimed to determine the influence of grouping feature harmonization and trait aggregation on trait-environment relationships. In particular, we aimed to (1) determine the effect of different harmonization methods using 4 invertebrate grouping feature datasets for seven grouping features from Europe, North America, New Zealand, and Australia. Furthermore, we (2) compared trait affinities obtained through different trait aggregation methods to trait affinities assigned at family-level by experts. We (3) re-analysed data on the effect of anthropogenic salinisation on biological traits by Szöcs et al. \cite{szocs_effects_2014} using harmonized grouping features and aggregated traits. By comparison with the original analysis, we investigated how harmonizing and aggregating trait data can modify the outcome of trait-environment relationships. Finally, we (4) present an overview of discrepancies in trait definition between the used invertebrate trait datasets and discuss challenges of trait data synthesis.

@@

We extracted information from 6 trait databases from Europe, North America, Australia, and New Zealand and harmonized 7 grouping features. Trait information for Europe was obtained from the freshwaterecology.info database \cite{schmidt-kloiber_www.freshwaterecology.info_2015} and the Tachet database \cite{usseglio-polatera_biomonitoring_2000}. The freshwaterecology.info contains taxa on species-level, while taxa recorded in the Tachet database are on species, genus and family-level.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% freshwater.info requested from ASK
% Slightly changed - Comment from ASK: this is rather confusing at this stage as most of the assignments in freshwaterecology.info (fwe) are on species level whereas Tachet et al. have mostly genus and family level data in their database
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Trait information for North America was obtained from Twardochleb et al. \cite{twardochleb_trait_data_2020} and complemented by Vieira et al. \cite{vieira_database_nodate}. Data on body form for European and North American taxa was provided based on expert knowledge \cite{polatera_personal_information_2020}. For Australia and New Zealand, we used trait databases from Kefford et al. \cite{kefford_integrated_2020} and Philips and Smith respectively \cite{Philips_and_Smith_NZ_DB_2018}. To increase readability we refer to the databases as well as the datasets we extracted from them by the name of the continent they originate from. Excluded from this are the European databases, which we refer to by their commonly used names (freshwaterecology.info and Tachet database). 
% Eingefügt da die Bennung uneinheitlich im Manuskript war, auch von den Co-Autoren kritisiert wurde

We selected traits of seven grouping features that were available in all databases, are commonly used, and describe different parts of the biology of a species: life history (Voltinism), morphology (Respiration, Body form, Size), ecology (Locomotion, Feeding mode) and reproduction (Oviposition). We omitted ecological traits that describe habitat preferences (e.g. temperature preference) because these traits are missing in the New Zealand trait database. The grouping features were differently classified across the databases, we therefore harmonized them into 26 traits (Table \ref{tab:traits_harmonization}). Harmonization was undertaken by amalgamating similar traits into one trait (e.g. crawlers and sprawlers into crawlers).
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Does it mean that the relative frequency of „non-amalgamated“ traits could be modified in the „grouping feature“ profiles of taxa? (because the highest score could be lower than the sum of the affinity scores of the amalgamated traits)
Thereby, for a particular taxa the highest trait affinity score among the amalgamated traits was taken.

We used fuzzy coded traits for establishing our harmonized datasets unless data quality prohibited. In the latter case we used binary traits, i.e. categorical and continuous traits were converted into binary traits. For example, in the  freshwaterecology.info database taxa were classified according to different floristic regions for the trait voltinism. Hence, the entries of the voltinism traits such as "arctic" or "boreal" were substituted with a value of 1. Implicitly, we assumed for binary traits that a value of 1 corresponds to the highest affinity for a particular trait and 0 to no affinity for a particular trait. Fuzzy codes are reported with different ranges in the trait databases (e.g. freshwaterecology.info 0 to 10, Tachet database 0 to 3 or 0 to 5). We standardized them to a range between 0 and 1 and converted trait affinities to percentages. Thus, fuzzy coded and binary traits were in the same range.

Prior to harmonization we consolidated duplicate taxa on species, genus or family-level if present, by either applying the median for fuzzy coded traits, or the maximum for binary traits. We omitted taxa with a less precise taxonomic resolution than family-level.

\begin{table}[H]
\centering
\caption{Traits of the harmonized grouping features. The last column indicates traits that were amalgamated for harmonization (no amalgamation needed if empty).}
\label{tab:traits_harmonization}
\begin{tabular}{lll}
\toprule[.1em]
Grouping feature & Trait & Amalgamated traits\\
\toprule[.1em]
Voltinism    & \begin{tabular}[c]{@{}l@{}}Semivoltine\\ Univoltine\\ Bi/multivoltine\end{tabular}                                & \begin{tabular}[c]{@{}l@{}}\textless 1 generation per year\\ 1 generation per year\\ \textgreater 1 generation per year\end{tabular}                                                                                                                                            \\
\midrule
Body Form    & \begin{tabular}[c]{@{}l@{}}Cylindrical \\ Flattenend\\ Spherical\\ Streamlined\end{tabular}                       & \begin{tabular}[c]{@{}l@{}}Cylindrical, tubular\\ Flattend, dorsoventrally flattened \textsuperscript{\textit{a}}\\ Spherical, round (humped)\\ Streamlined, fusiform\end{tabular}                                                                                                                                                                          \\
\midrule
Size         & \begin{tabular}[c]{@{}l@{}}Small \\ Medium \\ Large\end{tabular}                                                  & \begin{tabular}[c]{@{}l@{}}\textless 9 mm, \textless 10 mm \textsuperscript{\textit{b}}\\ 9 - 16 mm, 10 - 20 mm\\ \textgreater 16 mm, \textgreater 20 mm\end{tabular}                                                                                                                \\
\midrule
Respiration  & \begin{tabular}[c]{@{}l@{}}Gills\\ Plastron/Spiracle\\ \\ \\ Tegument\end{tabular}                             & \begin{tabular}[c]{@{}l@{}}Temporary air store, tracheal gills, gills\\ Spiracular gills, atmospheric breathers, \\ plant breathers, functional spiracles, \\ air (plants), aerial, plastron/spiracle\\ Cutaneous, tegument \end{tabular}                                                                         \\
\midrule
Locomotion   & \begin{tabular}[c]{@{}l@{}}Burrower\\ Crawler\\ Sessil\\ Swimmer \end{tabular}                                     & \begin{tabular}[c]{@{}l@{}}Interstitial, boring, burrowing\\ Sprawler, walking, climber, clinger, crawler\\ Attached, sessil\\ Skating, diving, planctonic, swimming\end{tabular}                                                                                                                 \\
\midrule
Feeding mode & \begin{tabular}[c]{@{}l@{}}Filterer\\ \\ Gatherer\\ \\ Herbivore\\ \\ Parasite\\ Predator\\ Shredder \\ \\ \end{tabular} & \begin{tabular}[c]{@{}l@{}}Active/passive filterer, absorber, \\ filter-feeder, collector-filterer, filterer\\ Deposit-feeder, collector-gatherer, \\ detrivore, gatherer\\ Grazer, scraper, piercer herbivore, \\ herbivore, algal piercer, piercer (plants)\textsuperscript{\textit{c}}\\ \\ Piercer (animals)\textsuperscript{\textit{c}}, predator \\ Miner, xylophagus, shredder, \\ shredder detrivore\end{tabular} \\
\hline
Oviposition  & \begin{tabular}[c]{@{}l@{}}Aquatic eggs\\ \\ Ovoviviparity\\ Terrestrial eggs\end{tabular}                            & \begin{tabular}[c]{@{}l@{}}Eggs attached to substrate/plants/stones,\\ free/fixed eggs/clutches\\ \\ Terrestrial clutches, terrestrial \end{tabular}     \\
\bottomrule[.1em]
\end{tabular}
\end{table}
\begin{minipage}{\linewidth}{\fontsize{8}{10}\selectfont
    \textit{a} The trait "bluff (blocky)" occurred in the Vieira et al. \cite{vieira_database_nodate} database and was newly classified by expert knowledge into cylindrical and flattened \cite{polatera_personal_information_2020}. \\  
    \textit{b} Reflects the different size classifications by the North American trait databases and the other trait databases. \\
    \textit{c} The trait piercer was defined in the Tachet database for piercing plants and animals, in contrast to the other databases \cite{usseglio-polatera_biomonitoring_2000}. Taxa exhibiting this trait have been assigned to predators or herbivores based on expert knowledge \cite{polatera_personal_information_piercer_2020}.
}
\end{minipage}

@@

Traits of the harmonized grouping feature datasets were aggregated to family-level using three approaches. I) Direct aggregation of taxa to family-level giving equal weight to every taxon using the mean or median, denoted \textit{direct\_agg\textsubscript{mean}} and \textit{direct\_agg\textsubscript{median}}, respectively. II) Stepwise aggregation, i.e. first to the genus-level and subsequently to the family-level using the mean or median. This approach gives equal weights to each genus. Hereafter, we denote this aggregation type as \textit{stepwise\_agg\textsubscript{mean}} or \textit{stepwise\_agg\textsubscript{median}}, respectively. III) Aggregation using a weighted mean approach, denoted as \textit{weighted\_agg}. This method weights the genera according to the number of their species in the trait datasets regardless if for every used grouping feature information was available. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% PUP: Is it the „potential number of species“ (i.e. the number of species potentially present at the selected spatial scale) or the number of species with traits/grouping features described? The grouping feature information can be available for a different set of species from a grouping feature to another. Does it mean that genus weights can be different in the aggregated family profile according to the information available on traits (regardless of the actual number of species in these different genera?
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{figure}
  \centering
  \subfile{Flowchart/Flowchart_methods.tex}
  \caption{Data processing steps of the selected traits. Intermediate (gray) and main (orange) steps of data preparation are depicted. The dashed bottom box illustrates the different trait aggregation methods using a small made-up example (data in the upper left corner). The aggregation methods (blue) and intermediate steps of the aggregation methods (purple) are displayed. Abbreviations: EU: Europe, NOA: North America, AUS: Australia, NZ: New Zealand, DB: Database.}
\end{figure}
